sims

### 2024-11-05 sims/sims.Rmd
update sims.Rmd to create 2 new Figures (2 and 3 in ms.)

>>>>

# Section 11: Simulate hazard data for 1 subject, 1 condition, 200 trials, and plot empirical hazard and survivor functions with error bars (Figure 2)

## Simulate hazard data based on user-defined hazard function

## apply function sim_haz
data_single <- sim_haz(subj_n = 1, rep_n = 200) %>%
               filter(condition == "cond1") %>%
               mutate(period = bin,
                      trial = str_sub(rep, start=6, end=6)) %>%
               select(-c("contrast","cumsum","outcomel","outcomen","bin","rep")) 
head(data_single, n=20)
```

## set up life table

```{r}
lifetable <- data_single %>%
    mutate(event = str_c("event", event)) %>%
    group_by(condition,period) %>% 
    count(event) %>% 
    ungroup() %>% 
    pivot_wider(names_from = event,
                values_from = n) %>% 
    mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
           event1 = ifelse(is.na(event1),0,event1),
           risk_set = event0 + event1) %>% # define the risk set
    mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
    mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4), # standard error for hazard
           pmass = event1/max(risk_set),
           se_pmass = sqrt((pmass * (1 - pmass)) / max(risk_set))) %>% 
    group_by(condition) %>%
    mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
           term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
           se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
    ungroup() 
```

## wrangle life table for plotting

```{r}
minimallt <- lifetable %>% select(period,condition, hazard, se_haz, survival, se_surv) %>%
  pivot_longer(cols = c(hazard,survival),
               names_to = "dv",
               values_to = "probability") %>%
  mutate(se = ifelse(dv=="hazard", se_haz, se_surv))
```

## plot hazard and survivor functions

```{r}
p11.1 <-ggplot(minimallt, aes(x = period, y = probability,
                         colour = dv)) +
  geom_line(linewidth = 1) +
  geom_point(size=3) +
  geom_linerange(aes(ymin=probability-se, ymax=probability+se), show.legend = F) +
  scale_colour_brewer(palette = "Dark2") + 
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0,1)) +
  ggtitle("hazard and survivor functions") +
  xlab("timebin")
p11.1

ggsave("sims/figures/haz_surv_single.jpeg",
       width = 7, height = 5, dpi=800)
```

# Section 12 (Figure 3): plot of hazard and ca(t) for 2 conditions with error bars

## hazard

### generate data set

```{r}
## make it reproducible
set.seed(8)

## apply function sim_haz
data_single2 <- sim_haz(subj_n = 1, rep_n = 2000) %>%
               mutate(period = bin,
                      trial = str_sub(rep, start=6, end=6)) %>%
               select(-c("contrast","cumsum","outcomel","outcomen","bin","rep")) 
head(data_single2, n=20)
```

### set up life table

```{r}
lifetable2 <- data_single2 %>%
    mutate(event = str_c("event", event)) %>%
    group_by(condition,period) %>% 
    count(event) %>% 
    ungroup() %>% 
    pivot_wider(names_from = event,
                values_from = n) %>% 
    mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
           event1 = ifelse(is.na(event1),0,event1),
           risk_set = event0 + event1) %>% # define the risk set
    mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
    mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4), # standard error for hazard
           pmass = event1/max(risk_set),
           se_pmass = sqrt((pmass * (1 - pmass)) / max(risk_set))) %>% 
    group_by(condition) %>%
    mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
           term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
           se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
    ungroup() 
```

### plot hazard functions

```{r}
p12.1 <-ggplot(lifetable2, aes(x = period, y = hazard,
                         colour = condition)) +
  geom_line(linewidth = 1) +
  geom_point(size=3) +
  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
  scale_colour_brewer(palette = "Dark2") + 
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0,1)) +
  ggtitle("hazard")  +
  xlab("timebin") + ylab("probability")
p12.1
```

## accuracy

```{r}
acc <- tibble(
  time = seq(1,10,1),
  cond1 = rep(1,length = 10),
  cond2 = c(0, 0, 0, 0, 0, 1, 1, 1, 1, 1)
) %>% 
  pivot_longer(-time,
               names_to = "condition",
               values_to = "accuracy") %>% 
  mutate(condition = factor(condition))
head(acc,n=20)
```

```{r}
p12.2 <-ggplot(acc, aes(x = time, y = accuracy,
                         colour = condition)) +
  geom_line(linewidth = 1) +
  geom_point(size=3) +
  scale_colour_brewer(palette = "Dark2") + 
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(limits = c(0,1)) +
  xlab("timebin") +
  ylab("probability") +
  ggtitle("conditional accuracy") 
p12.2
```

Combine plots

```{r}
p12.3 <- (p12.1 / p12.2) +
  plot_layout(guides = 'collect',
              axes = 'collect')
p12.3

ggsave("sims/figures/haz_acc_single.jpeg",
       width = 7, height = 5, dpi=800)

```
